---
title: "Thesis"
author: "Thiyaghessan"
date: "2023-04-30"
output:
  pdf_document:
    latex_engine: xelatex
  html_document: default
---

## Analysis Replication Guide

This markdown file replicates all plots produced in the accompanying thesis.

## Load packages and configure environment

```{r, echo=FALSE, setup}
rm(list = ls()) # Clear Environment
# Set working directory to where the data files are located
knitr::opts_knit$set(root.dir = "C:/Users/thiya/OneDrive/Uchicago/thesis/data")

library(dplyr)
library(tidyverse)
library(igraph)
library(kableExtra)
library(intergraph)
library(ergm)
library(texreg)
library(stringr)

windowsFonts("Arial" = windowsFont("Arial")) # Set font for plots
```

## Global Variables

```{r}
# These podcasts are far away from the center of the network and are excluded from visualization
Extra_Exclude <- c("POLITICO Energy", "Washington Today")

# Set color pallette for network visualizations
Viz_Colors <- c("darkred", "tomato", "gray80", "skyblue", "darkblue", "gray80")

# Configurations for Degree Distribution Graphs
Degree_Ideologies <- c("Reactionary", "Conservative", "Moderate", "Liberal", "Radical", "NULL")
Degree_Colors <- c("tomato", "skyblue", "gray50", "black", "darkblue", "darkred")

# Configurations for Betweenness Centrality Distribution Plot
Btwn_Ideologies <- c("Conservative", "Liberal", "Moderate", "Radical", "Reactionary", "NULL")
Btwn_Colors <- c("tomato", "skyblue", "gray50", "darkblue", "darkred", "black")

# Configurations for Power Law Ideologies
Plaw_Ideologies <- c("Reactionary", "Conservative", "Moderate", "Liberal", "Radical")
```


## Required Functions

```{r}
# Function to create a dataframe that aggregates degree distribution by ideology
make_deg_df <- function(g, mode, ideologies){
 for (i in 1:6){
   deg_ideology <- igraph::degree(g,
                                  v = V(g)$ideology == i,
                                  mode = mode)
   deg_df <- as.data.frame(table(deg_ideology)) %>%
     mutate(ideology = ideologies[i])
   if (i == 1){
     master_deg_df <- deg_df
     } else {
       master_deg_df <- rbind(master_deg_df, deg_df)    
     }
 }
  return(master_deg_df)
}

# Function to create a dataframe that contains power law fits for degree distributions by ideology
make_deg_plaw_df <- function(ideologies, deg_df){
  for (id in ideologies){
  df = as.data.frame(power.law.fit(deg_df$Freq[deg_df$ideology == id]))
  if (id == "Reactionary"){
    plaw_df = df
  } else {
    plaw_df = rbind(plaw_df, df)
  }
  }
  return(plaw_df)
}

```


## Data Preprocessing

```{r}
# Import csv containing information on podcasts
pod_df <- read.csv("podcast_hosts.csv") # Data on podcasts
pod_bias_df <- read.csv("podcast_bias.csv") # Data on the level of Bias 
hosts_df <- read.csv("podcast_hostattribs.csv") # Data on the podcast hosts' occupations/source of fame

# Merge dataframes
nodes_df <- left_join(pod_df, pod_bias_df, by = "podcasts")

# Summarise the number of unique distributors present in the dataset
distrib_df <- nodes_df %>%
  group_by(parent) %>% 
  summarise(count = n())
# Isolate distributors who own/represent more than one podcast in the dataset
distributors <- distrib_df$parent[distrib_df$count > 1]

# Add this information to nodes_df and recode other variables accordingly
nodes_df <- nodes_df %>% 
  mutate(distributor = ifelse(parent %in% distributors, parent, 
                            ifelse(parent != "Independent", "Single Company", "Independent")),
         distributor_code = ifelse(distributor == "Independent", 3,
                                   ifelse(distributor == "Single Company", 2, 1)),
         bias_ratio = numbiased / (numbiased + numunbiased),
         bias_ratio = ifelse(is.nan(bias_ratio), mean(bias_ratio, na.rm = TRUE), bias_ratio),
         bias_ratio = ifelse(bias_ratio > 0, bias_ratio, mean(bias_ratio)),
         id_code = ifelse(main_ideology == "reactionary", 1, 
                          ifelse(main_ideology == "conservative", 2, 
                                 ifelse(main_ideology == "moderate", 3, 
                                        ifelse(main_ideology == "liberal", 4, 
                                               ifelse(main_ideology == "radical", 5, 6))))))


# Import data frame with podcast collaborations
collab_df <- read.csv("podcast_collabs.csv") %>% 
  rename("from" = "From",
         "to" = "To") %>% 
  group_by(from, to) %>% 
  summarise(weight = n()) # weight edges by number of repeat appearances

# Import data summarizing total number of cross-partisan/homophilic collaborations
crosspart_df <- read.csv("crosspart2.csv") %>% 
                group_by(crosspart) %>% 
                summarise(count = sum(count),
                          weight = sum(weight),
                          homophilic = mean(homophilic),
                          cross = mean(cross))

# create social network graph
g_full <- graph_from_data_frame(collab_df,
                           directed = TRUE,
                           vertices = nodes_df$podcasts) # Directed edges

# Set node attributes
g_full <- g_full %>% 
  set_vertex_attr("ideology",
                  index = V(g_full),
                  value = nodes_df$id_code) %>% 
  set_vertex_attr("distributor",
                  index = V(g_full),
                  value = nodes_df$distributor_code) %>% 
  set_vertex_attr("bias_count",
                  index = V(g_full),
                  value = nodes_df$numbiased) %>%
  set_vertex_attr("bias_ratio",
                  index = V(g_full),
                  value = nodes_df$bias_ratio) 
```


## Data Preprocessing for Visualization

The first step is to load in the necessary data and functions

```{r, echo=FALSE}
# We want to exclude nodes in the visualization that have no edges.
# If we don't do that, R will zoom  out and make the plot very small
main_exclude <- unique(c(collab_df$from, collab_df$to, Extra_Exclude))

# Subset the data to only include nodes and edges for visualization
nodes_viz_df <- nodes_df %>% 
  filter(podcasts %in% main_exclude[! main_exclude %in% Extra_Exclude])

collab_viz_df <- collab_df %>% 
  filter(from %in% main_exclude[! main_exclude %in% Extra_Exclude],
         to %in% main_exclude[! main_exclude %in% Extra_Exclude])

# create social network graph for visualization
g_viz <- graph_from_data_frame(collab_viz_df,
                               directed = TRUE,
                               vertices = nodes_viz_df$podcasts) # Directed edges

# Set node attributes
g_viz <- g_viz %>% 
  set_vertex_attr("ideology",
                  index = V(g_viz),
                  value = nodes_viz_df$id_code)

# Set node color based on ideology
V(g_viz)$color <- Viz_Colors[V(g_viz)$ideology] 

# Scale node size based on the number of collaborations
V(g_viz)$size <- log(igraph::degree(g_viz, mode = "all"))

# Set edge attributes

# Set arrow size
E(g_viz)$arrow.size <- 0.01

# Scale edges on log-scale according to weight
E(g_viz)$width <- log10(E(g_viz)$weight)

# Assign degree to each node
deg <- igraph::degree(g_viz, mode = "all")

# Set layout
l <- layout_with_kk(g_viz)
```


## Network Visualization

```{r}
par(mar=c(0,0,0,0)+1)
plot(g_viz,
     vertex.label = ifelse(deg[V(g_viz)$name] > 100, V(g_viz)$name, NA), # Display labels for podcasts with degree > 100
     vertex.label.family = "Arial",
     vertex.label.color = "black",
     vertex.label.cex = 0.5,
     edge.color = "gray90",
     main = "Collaborations Among The 250 Most Popular Political Podcasts",
     margin = 0)

legend(
  "bottomright",
  legend = c("Reactionary", "Conservative", "Moderate", "Liberal", "Radical"),
  pt.bg  = c("darkred", "tomato", "gray50", "skyblue", "darkblue"),
  pt.cex = c(2, 2, 2),
  pch    = 21,
  cex    = 0.8,
  bty    = "o",
  title  = "Ideological Affiliation"
)

legend(
  "bottomleft",
  legend = c("< 10", "10 - 20", "20 - 40", "40 - 60", "> 60"),
  pt.bg  = "gray50",
  pt.cex = c(0.4, 0.8, 1.2, 1.6, 1.8),
  pch    = 21,
  cex    = 0.8,
  bty    = "o",
  title  = "Number of Collaborations"
)
```

## Indegree Distribution Plot

```{r}
# Create DataFrame with distribution
indeg_df <- make_deg_df(g_full, "in", Degree_Ideologies)
# Reorder factor levels
indeg_df$deg_ideology <- factor(indeg_df$deg_ideology,
                                levels = c("0", "1", "2", "3", "4", "5", "6", "7", 
                                           "8", "9" , "10", "11", "12", "13", "14",
                                           "15", "16", "17", "19", "24", "26", "27", "28"))
# Bar Plot
indeg_dist_barplot <- ggplot(indeg_df,
                             aes(x = deg_ideology,
                                 y = Freq,
                                 fill = ideology)) +
                        geom_bar(stat = "identity",
                                 position = "stack") +
                        xlab("Number of Guests Hosted") +
                        ylab("Number of Podcasts") +
                        labs(title = "In Degree Distribution by Ideology") +
                        scale_fill_manual(name = "Ideology",
                                          values = Degree_Colors) +
                        theme_bw(base_size = 18) +
                        theme(plot.title = element_text(hjust = 0.5))
indeg_dist_barplot
# Compute mean and median
sprintf("The Mean of this distribution is %.2f", mean(as.integer(indeg_df$deg_ideology)))
sprintf("The Median of this distribution is %.2f", median(as.integer(indeg_df$deg_ideology)))
```

## Outdegree Distribution Plot

```{r}
# Create DataFrame with distribution
outdeg_df <- make_deg_df(g_full, "out", Degree_Ideologies)
# Reorder factor levels
outdeg_df$deg_ideology <- factor(outdeg_df$deg_ideology,
                                levels = c("0", "1", "2", "3", "4", "5", "6", "7", 
                                           "8", "9" , "10", "11", "12", "13","14","16", "17", 
                                           "18", "21", "22", "25", "27", "31", "45", "63", "67"))
# Bar Plot
outdeg_dist_barplot <- ggplot(outdeg_df,
                              aes(x = deg_ideology,
                                  y = Freq,
                                 fill = ideology)) +
                       geom_bar(stat = "identity",
                                position = "stack") +
                       xlab("Number of Guests Hosted") +
                       ylab("Number of Podcasts") +
                       labs(title = "In Degree Distribution by Ideology") +
                       scale_fill_manual(name = "Ideology",
                                         values = Degree_Colors) +
                       theme_bw(base_size = 18) +
                       theme(plot.title = element_text(hjust = 0.5))
outdeg_dist_barplot
# Compute mean and median
sprintf("The Mean of this distribution is %.2f", mean(as.integer(outdeg_df$deg_ideology)))
sprintf("The Median of this distribution is %.2f", median(as.integer(outdeg_df$deg_ideology)))
```

## Betweenness Centrality Plot

```{r}
# Betweenness centrality

# Calculate betweenness centrality
btwn <- igraph::betweenness(
  g_full,
  v = V(g_full),
  directed = TRUE,
  weights = E(g_full)$weight,
  nobigint = TRUE,
  normalized = TRUE,
  cutoff = -1 # remove those with negative betweenness (these are anomalous)
)

# Store betweenness centrality into a dataframe
btwn_df <- as.data.frame(btwn) %>% 
  rownames_to_column() %>% 
  rename(podcasts = rowname) %>% 
  left_join(nodes_df, by = "podcasts") %>% 
  mutate(ideology = as.factor(main_ideology)) %>% 
  filter(btwn > 0.0001)

# Plot distribution
btwn_plot <- ggplot(btwn_df,
             aes(x = reorder(podcasts, -btwn),
                 y = btwn,
                 fill = main_ideology)) +
             geom_bar(stat = "identity")  +
             xlab("Podcast Names") +
             ylab("Normalized Betweenness Centrality") +
             labs(title = "Podcast Betweenness Centrality Distribution by Ideology") +
             scale_fill_manual(name = "Ideology",
                               values = Btwn_Colors,
                               labels = Btwn_Ideologies) +
             theme_bw(base_size = 18) +
             theme(plot.title = element_text(hjust = 0.5)) +
             theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6))
btwn_plot

```

## Distribution of Unique Collaborations

```{r}
crosspart_uniq_plot <- ggplot(crosspart_df,
                              aes(x = reorder(crosspart, -count), 
                                  y = count, 
                                  fill = cross)) +
                      geom_bar(stat = "identity") +
                      labs(title = "Distribution of Unique Cross Partisan Collaborations",
                           subtitle = "Color Scaled According to Magnitude of Ideological Difference") +
                      xlab("Type of Collaboration") + 
                      ylab("Number of Unique Collaborations") +
                      scale_fill_gradient(name = "Cross Partisan Rating \n (4 = High, 0 = Low)",
                                          low = "darkred", high = "grey")+
                      scale_x_discrete(labels = function(x) str_wrap(x, width = 20)) +
                      theme_bw(base_size = 18) +
                      theme(plot.title = element_text(hjust = 0.5)) +
                      theme(plot.subtitle = element_text(hjust = 0.5)) +
                      theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18))
crosspart_uniq_plot
```

## Distribution of All Collaborations

```{r}
crosspart_totl_plot <- ggplot(crosspart_df,
                              aes(x = reorder(crosspart, -weight), 
                                  y = weight,
                                  fill = cross)) +
                      geom_bar(stat = "identity") +
                      labs(title = "Distribution of Unique Cross Partisan Collaborations",
                           subtitle = "Color Scaled According to Magnitude of Ideological Difference") +
                      xlab("Type of Collaboration") + 
                      ylab("Number of Unique Collaborations") +
                      scale_fill_gradient(name = "Cross Partisan Rating \n (4 = High, 0 = Low)",
                                          low = "darkred", high = "grey")+
                      scale_x_discrete(labels = function(x) str_wrap(x, width = 20)) +
                      theme_bw(base_size = 18) +
                      theme(plot.title = element_text(hjust = 0.5)) +
                      theme(plot.subtitle = element_text(hjust = 0.5)) +
                      theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18))
crosspart_totl_plot
```

## Power Law Fits

```{r}
# Indegree distibution
indeg_plaw_df <- make_deg_plaw_df(Plaw_Ideologies, indeg_df)
# Create table visualization
indeg_plaw_df %>% 
  rename(Fit = continuous) %>% 
  mutate(Fit = Plaw_Ideologies) %>% 
  kbl(caption = "Power Law Fit on In Degree Distribution") %>%
  kable_classic(full_width = F, html_font = "Garamond")
```

```{r}
# Out degree Distribution
outdeg_plaw_df <- make_deg_plaw_df(Plaw_Ideologies, outdeg_df)
# Create table visualization
outdeg_plaw_df %>% 
  rename(Fit = continuous) %>% 
  mutate(Fit = Plaw_Ideologies) %>% 
  kbl(caption = "Power Law Fit on Out Degree Distribution") %>%
  kable_classic(full_width = F, html_font = "Garamond")
```

```{r}
# Betweenness centrality Power Law fit 
as.data.frame(power.law.fit(btwn_df$btwn)) %>% 
  rename(Fit = continuous) %>% 
  mutate(Fit = c("Betweenness")) %>% 
  kbl(caption = "Power Law Fit on Betweenness Centrality Distribution") %>%
  kable_classic(full_width = F, html_font = "Garamond")
```

```{r}
# Unique Collaborations Power Law Fit
as.data.frame(power.law.fit(crosspart_df$count)) %>% 
  rename(Fit = continuous) %>% 
  mutate(Fit = c("Betweenness")) %>% 
  kbl(caption = "Power Law Fit on Unique Cross-Partisan Collaborations") %>%
  kable_classic(full_width = F, html_font = "Garamond")
```

```{r}
# Total Collaborations Power Law Fit
as.data.frame(power.law.fit(crosspart_df$weight)) %>% 
  rename(Fit = continuous) %>% 
  mutate(Fit = c("Betweenness")) %>% 
  kbl(caption = "Power Law Fit on All Cross-Partisan Collaborations") %>%
  kable_classic(full_width = F, html_font = "Garamond")
```

## ERGM Models

### Data Preprocessing

```{r}
# Create dataframe for edges in ergm.
# Exclude collaboraitons involving "NULL" ideologies (remove NULL values since they are too sparse for simulations)

# Map ideologies to podcasts involved in collaborations
ideology_df <- nodes_df %>% 
  select(podcasts, main_ideology)

# First merge to get ideology of hosts
collab_ergm_df <- collab_df %>% 
  merge(x = collab_df,
        y = ideology_df,
        by.x = "to",
        by.y = "podcasts") %>% 
  rename(ideology_host = main_ideology)

# Second merge to get ideology of guests and filter out "NULL" ideologies
collab_ergm_df <- collab_ergm_df %>% 
  merge(x = collab_ergm_df,
        y = ideology_df,
        by.x = "from",
        by.y = "podcasts") %>% 
  rename(ideology_guest = main_ideology) %>% 
  filter(! ideology_host %in% "NULL",
         ! ideology_guest %in% "NULL") %>% 
  mutate(crosspart = str_c(ideology_guest, " - ", ideology_host)) %>% 
  select(from, to, weight)



# Load in additional data on host bios
hostbio_df <- read.csv("podcast_hostattribs.csv") %>%
  rename(podcasts = Podcast,
         alternative_media_personality = other,
         political_personality = politician,
         professional_personality = political_professional,
         academic_personality = academic) %>% 
  mutate(legacy_media_personality = tv_personality + print_personality + radio_personality) %>% 
  select(2, 4, 8:12) %>% 
  mutate(sum = rowSums(across(where(is.numeric)))) %>% 
  mutate(across(2:7, ~ . / sum)) %>% 
  mutate_at(2:7, ~replace(., is.nan(.), 0)) %>% 
  select(1:7)

# Subset nodes_df to exclude "NULL" Values
nodes_ergm_df <- nodes_df %>% 
  left_join(hostbio_df, by = "podcasts") %>% 
  filter(! main_ideology %in% "NULL") %>% 
  mutate_at(12:17, ~ifelse(. > 0, 1, 0))

# create social network graph
g_ergm <- graph_from_data_frame(collab_ergm_df,
                                directed = TRUE,
                                vertices = nodes_ergm_df$podcasts) # Directed edges

# Set node attributes
g_ergm <- g_ergm %>% 
  set_vertex_attr("ideology",
                  index = V(g_ergm),
                  value = nodes_ergm_df$id_code) %>% 
  set_vertex_attr("distributor",
                  index = V(g_ergm),
                  value = nodes_ergm_df$distributor_code) %>% 
  set_vertex_attr("bias_count",
                  index = V(g_ergm),
                  value = nodes_ergm_df$bias_std) %>%
  set_vertex_attr("bias_ratio",
                  index = V(g_ergm),
                  value = nodes_ergm_df$bias_ratio)  %>% 
  set_vertex_attr("political_personality",
                  index = V(g_ergm),
                  value = nodes_ergm_df$political_personality)  %>%  
  set_vertex_attr("academic_personality",
                  index = V(g_ergm),
                  value = nodes_ergm_df$academic_personality)  %>%
  set_vertex_attr("professional_personality",
                  index = V(g_ergm),
                  value = nodes_ergm_df$professional_personality)  %>%  
  set_vertex_attr("alt_media_personality",
                  index = V(g_ergm),
                  value = nodes_ergm_df$alternative_media_personality)  %>%  
  set_vertex_attr("religious_personality",
                  index = V(g_ergm),
                  value = nodes_ergm_df$religious_personality)  %>% 
  set_vertex_attr("legacy_media_personality",
                  index = V(g_ergm),
                  value = nodes_ergm_df$legacy_media_personality)

# Network object for ERGM
network <- intergraph::asNetwork(g_ergm)
```


### Model 1

#### Build Model

```{r, echo = FALSE}
# mod_1 <- ergm(network ~ edges +
#               nodematch("ideology", diff = T) +
#               nodeofactor("ideology") +
#               nodeifactor("ideology") +
#               nodecov("bias_ratio") +
#               nodecov("political_personality") +
#               nodecov("legacy_media_personality") +
#               nodecov("alt_media_personality") +
#               nodecov("religious_personality") +
#               nodecov("academic_personality") +
#               nodematch("distributor", diff = T) +
#               odegree(0:5) +
#               mutual,
#               control = control.ergm(MCMLE.maxit= 40))
# Save Model
# saveRDS(mod_1, file="mod_1.RData")
```

#### Model Results

```{r}
mod_1 <- readRDS("mod_1.RData")
table <- htmlreg(list(mod_1),
        custom.coef.names = c("Number of Edges",
                              "Reactionary",
                              "Conservative",
                              "Moderate",
                              "Liberal",
                              "Radical",
                              "Difference in In Degree - Conservative",
                              "Difference in In Degree - Moderate",
                              "Difference in In Degree - Liberal",
                              "Difference in In Degree - Radical",
                              "Difference in Out Degree - Conservative",
                              "Difference in Out Degree - Moderate",
                              "Difference in Out Degree - Liberal",
                              "Difference in Out Degree - Radical",
                              "Ratio of Biased to Unbiased Statements",
                              "Host is a political figure",
                              "Host is a legacy media figure",
                              "Host is a alternative media figure",
                              "Host is a relgious leader",
                              "Host is an academic",
                              "Independent Podcast",
                              "Podcast owned by a Single Company",
                              "Podcast part of a Distribution Network",
                              "0 Guest Appearances",
                              "1 Guest Appearances",
                              "2 Guest Appearances",
                              "3 Guest Appearances",
                              "4 Guest Appearances",
                              "5 Guest Appearances",
                              "Presence of Mutual Connection"),
        bold = 0.5)
htmltools::HTML(table)
```

#### Diagnostics

```{r}
gof(mod_1)
```
```{r}
plot(gof(mod_1))
```

### Model 2

#### Build model

```{r, echo=FALSE}
#Run and save model

# mod_2 <- ergm(network ~ edges +
#               nodemix("ideology", base = NULL) +
#               nodecov("bias_ratio") +
#               mutual,
#             control = control.ergm(MCMLE.maxit= 40))
# saveRDS(mod_2, file="mod_2.RData")
```

#### Model Results

```{r}
mod_2 <- readRDS("mod_2.RData")
table <- htmlreg(list(mod_2),
       custom.coef.names = c("Number of Edges",
                             "Conservative -> Reactionary",
                             "Moderate -> Reactionary",
                             "Liberal -> Reactionary",
                             "Radical -> Reactionary",
                             "Reactionary -> Conservative",
                             "Conservative -> Conservative",
                             "Moderate -> Conservative",
                             "Liberal -> Conservative",
                             "Radical -> Conservative",
                             "Reactionary -> Moderate",
                             "Conservative -> Moderate",
                             "Moderate -> Moderate",
                             "Liberal -> Moderate",
                             "Radical -> Moderate",
                             "Reactionary -> Liberal",
                             "Conservative -> Liberal",
                             "Moderate -> Liberal",
                             "Liberal -> Liberal",
                             "Radical -> Liberal",
                             "Reactionary -> Radical",
                             "Conservative -> Radical",
                             "Moderate -> Radical",
                             "Liberal -> Radical",
                             "Radical -> Radical",   
                             "Ratio of Biased to Unbiased Statements",
                             "Presence of Reciprocal Collaboration"),
       bold = 0.05)
htmltools::HTML(table)
```

#### Diagnostics

```{r}
gof(mod_2)
```
```{r}
plot(gof(mod_2))
```

